Large squeezing in Bose-Einstein condensates with large numbers of atoms

### 2013-07-17 large_number_squeezing_with_atoms.tex
Fix typo in Josephson paragraph.

When considering squeezing, the Hamiltonian (\ref{eqTwoModeHamiltonian}) has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}. This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
\begin{eqnarray}
\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b} \\
\hat{J}_- &=& \hat{a} \hat{b}^{\dagger} \\
\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right) \\
\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right) \\
\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
\end{eqnarray}
Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
\begin{equation}
\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2
\label{eqNonlinearJosephson}
\end{equation}
where $\hat{N}=\hat{n}_1 + \hat{n}_2$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.  In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
\begin{equation}
  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
\end{equation}
The system described by $\hat{H}_J$ is then solved using angular momentum algebra.

>>>>

When considering squeezing, the Hamiltonian (\ref{eqTwoModeHamiltonian}) has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}. This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
\begin{eqnarray}
\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b} \\
\hat{J}_- &=& \hat{a} \hat{b}^{\dagger} \\
\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right) \\
\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right) \\
\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
\end{eqnarray}
Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
\begin{equation}
\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2
\label{eqNonlinearJosephson}
\end{equation}
where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_1 - \chi_2\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.  In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
\begin{equation}
  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
\end{equation}
The system described by $\hat{H}_J$ is then solved using angular momentum algebra.
